---
title: "Solution 3"
subtitle: HUDM 4125
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**2.7-1**

(a) \[
P(3 \leq X \leq 5) = P(X = 3) + P(X = 4) + P(X = 5)
\]

\[
P(X = 3) = \frac{4^3 e^{-4}}{3!} = \frac{64 e^{-4}}{6}, \quad P(X = 4) = \frac{4^4 e^{-4}}{4!} = \frac{256 e^{-4}}{24}, \quad P(X = 5) = \frac{4^5 e^{-4}}{5!} = \frac{1024 e^{-4}}{120}
\]

Thus,

\[
P(3 \leq X \leq 5) = e^{-4} \left( \frac{64}{6} + \frac{256}{24} + \frac{1024}{120} \right) = 0.547
\]

(b) \[
P(X \geq 3) = 1 - P(X \leq 2)
\]

\[
P(X \leq 2) = P(X = 0) + P(X = 1) + P(X = 2)
\]

\[
P(X = 0) = e^{-4}, \quad P(X = 1) = 4e^{-4}, \quad P(X = 2) = 8e^{-4}
\]

\[
P(X \leq 2) = e^{-4} \times 13
\]

Thus,

\[
P(X \geq 3) = 1 - 13e^{-4} = 0.7619
\]

(c) \( P(X \leq 3) \)

\[
P(X \leq 3) = P(X = 0) + P(X = 1) + P(X = 2) + P(X = 3)
\]

\[
P(X \leq 3) = e^{-4} \left( 1 + 4 + 8 + \frac{64}{6} \right)
\]

Thus,

\[
P(X \leq 3) = e^{-4} \left( 13 + \frac{64}{6} \right) = 0.4335
\]


\vspace{1em}

**2.7-4**  

The probability mass function for a Poisson distribution is:

\[
P(X = k) = \frac{\lambda^k e^{-\lambda}}{k!}
\]

\[
3P(X = 1) = P(X = 2)
\]

Substitute the Poisson PMF for \( P(X = 1) \) and \( P(X = 2) \):

\[
3 \times \frac{\lambda^1 e^{-\lambda}}{1!} = \frac{\lambda^2 e^{-\lambda}}{2!}
\]

\[
3 \lambda e^{-\lambda} = \frac{\lambda^2 e^{-\lambda}}{2}
\]

\[
3 \lambda = \frac{\lambda^2}{2}
\]

\[
6 \lambda = \lambda^2
\]

\[
\lambda^2 - 6 \lambda = 0
\]

\[
\lambda(\lambda - 6) = 0
\]

Thus, \( \lambda = 6 \)

\[
P(X = 4) = \frac{6^4 e^{-6}}{4!} = \frac{1296 e^{-6}}{24}
\]

\[
P(X = 4) \approx 0.1339
\]

\vspace{1em}

**2.7-5** 
 
\[
\lambda = \frac{225}{150} = 1.5
\]

\[
P(X \leq 1) = P(X = 0) + P(X = 1)
\]

\[
P(X = 0) = \frac{1.5^0 e^{-1.5}}{0!} = e^{-1.5}
\]

\[
P(X = 1) = \frac{1.5^1 e^{-1.5}}{1!} = 1.5 e^{-1.5}
\]

\[
P(X \leq 1) = e^{-1.5} + 1.5 e^{-1.5} = 0.5578
\]

\vspace{1em}

**2.7-13** 

\[
\frac{\lambda^2 e^{-\lambda}}{2!} = 4 \times \frac{\lambda^3 e^{-\lambda}}{3!}
\]

\[
\frac{\lambda^2}{2} = 4 \times \frac{\lambda^3}{6}
\]

\[
3 \lambda^2 = 4 \lambda^3
\]

\[
\lambda = \frac{3}{4}
\]

\[
E(X^2) = \lambda + \lambda^2
\]

\[
E(X^2) = \frac{3}{4} + \left(\frac{3}{4}\right)^2 = 1.3125
\]

\vspace{1em}

**3.1-2** 

(a)  \[
\int_0^1 c(x - x^2) \, dx = 1
\]

\[
c = 6
\]

(b) \[
P(0.3 < X < 0.6) = \int_{0.3}^{0.6} 6(x - x^2) \, dx \approx 0.4320
\]

\vspace{1em}

**3.1-5** 

(a) \[
W = a + (b - a)Y
\]

The CDF of \( W \) is:

\[
F_W(w) = P(W \leq w) = P\left(Y \leq \frac{w - a}{b - a}\right)
\]

Since \( Y \sim U(0, 1) \), its CDF is:

\[
F_Y(y) = y
\]

Thus:

\[
F_W(w) = 
\begin{cases}
0, & w < a \\
\frac{w - a}{b - a}, & a \leq w \leq b \\
1, & w > b
\end{cases}
\]

(b) The random variable \( W \) is uniformly distributed between \( a \) and \( b \):

\[
W \sim U(a, b)
\]

\vspace{1em}

**3.1-9** 

(i) \[
P(0 \leq X \leq 1/2) = \int_0^{1/2} 2(1 - x) \, dx
\]

\[
P(0 \leq X \leq 1/2) = 0.75
\]

(ii) \[
P(1/4 \leq X \leq 3/4) = \int_{1/4}^{3/4} 2(1 - x) \, dx
\]

\[
P(1/4 \leq X \leq 3/4) = 0.5
\]

(iii) \[
P(X = 3/4) = 0
\]

(iv) \[
P(X \geq 3/4) = \int_{3/4}^{1} 2(1 - x) \, dx
\]

\[
P(X \geq 3/4) = 0.0625
\]

\vspace{1em}

**3.2-1** 

(a) 
The pdf of \( X \) is:

\[
f(x) =  \frac{1}{3}e^{-\frac{1}{3}x}, \quad x \geq 0
\]

For an exponential distribution, the mean is \( \frac{1}{\lambda} \). Therefore:

\[
\text{Mean} = 3
\]

The variance of an exponential distribution is \( \frac{1}{\lambda^2} \). Therefore:

\[
\text{Variance} = 9
\]

(b) This is also the MGF for an exponential distribution with rate \( \lambda = 3 \).   

The pdf of \( X \) is:

\[
f(x) = 3e^{-3x}, \quad x \geq 0
\]

\[
\text{Mean} = 0.3333
\]

\[
\text{Variance} = 0.1111
\]

\vspace{1em}

**3.2-13** 

(a) \[
P(14.85 < X < 32.01) = F(32.01) - F(14.85)
\]

Using the chi-square cumulative distribution function (CDF):

\[
P(14.85 < X < 32.01) \approx F_{23}(32.01) - F_{23}(14.85) = 0.9 - 0.1 = 0.8
\]

Thus:

\[
P(14.85 < X < 32.01) = 0.8
\]

(b) We use the inverse CDF (quantile function):
  
\[
a = \chi^2_{0.025}(23) = 11.69
\]

We use the quantile function:
  
\[
b = \chi^2_{0.975}(23) = 38.08
\]

Thus:

\[
P(a < X < b) = P(11.69 < X < 38.08) = 0.95
\]

(c) For a chi-square distribution with \( k \) degrees of freedom:

- The mean is \( k \), and
- The variance is \( 2k \).

Thus, for \( X \sim \chi^2(23) \):

\[
\text{Mean} = 23
\]
\[
\text{Variance} = 46
\]

(d) \[
\chi^2_{0.05}(23) = 35.17
\]
\[
\chi^2_{0.95}(23) = 13.09
\]

\vspace{1em}

**3.2-16** 

\[
P(T > 26.30) = 1 - F(26.30)
\]

Using R, we calculate:

```{r}
# Gamma distribution parameters
k <- 8
lambda <- 0.5

# Calculate the probability
prob <- 1 - pgamma(26.30, shape = k, rate = lambda)
round(prob, 4)
```

\vspace{1em}

**3.2-17** 
```{r}
# Chi-square distribution parameter
df <- 4

# Find P(X > 7.779)
p_chisq <- 1 - pchisq(7.779, df = df)

# Binomial distribution parameters
n <- 15
p <- p_chisq

# Find P(Y <= 3)
prob <- pbinom(3, size = n, prob = p)
round(prob, 4)
```